### 2024-03-01 main.tex
Create paper template and add a summary of NN-descent

\documentclass[sigconf, nonacm]{acmart}

\begin{document}
\title{Efficient Shapelet Computation for Silhouette Score}

\author{Adam Charane}
\affiliation{%
	\institution{Free University of Bozen-Bolzano}
	\city{Bolzano}
	\state{Italy}
}
\email{acharane@unibz.it}

\author{Matteo Ceccarello}
\affiliation{%
	\institution{University of Padova}
	\city{Padova}
	\state{Italy}
}
\email{matteo.ceccarello@dei.unipd.it}

\author{Johann Gamper}
\affiliation{%
	\institution{Free University of Bozen-Bolzano}
	\city{Bolzano}
	\state{Italy}
}
\email{johann.gamper@unibz.it}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
	Praesent imperdiet, lacus nec varius placerat, est ex eleifend justo, a vulputate leo massa consectetur nunc. Donec posuere in mi ut tempus. Pellentesque sem odio, faucibus non mi in, laoreet maximus arcu. In hac habitasse platea dictumst. Nunc euismod neque eu urna accumsan, vitae vehicula metus tincidunt. Maecenas congue tortor nec varius pellentesque. Pellentesque bibendum libero ac dignissim euismod. Aliquam justo ante, pretium vel mollis sed, consectetur accumsan nibh. Nulla sit amet sollicitudin est. Etiam ullamcorper diam a sapien lacinia faucibus.
\end{abstract}

\maketitle

\section{Introduction}

\section{Related Work}
NN-Descent~\cite{NN-Descent} is based on the principle: \textit{A neighbor of
a neighbor is likely to be a neighbor}. The idea is, given an approximation of 
the nearest neighbor, then the approximation can be improved by exploring the 
neighbors of the neighbors of each point. To improve the running time, the 
authors use local join to avoid comparing same points many times in the same 
iteration. They also use incremental search to avoid comparing points that have
been compared in previous iterations. Also, sampled is used to lower the 
overhead of local joins. Finally, random projection trees are used to 
initialize the graph.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}
\endinput

### 2024-03-01 main.tex
minor change

@@
 NN-Descent~\cite{NN-Descent} is based on the principle: \textit{A neighbor of
-a neighbor is likely to be a neighbor}. The idea is, given an approximation of 
-the nearest neighbor, then the approximation can be improved by exploring the 
-neighbors of the neighbors of each point. To improve the running time, the 
-authors use local join to avoid comparing same points many times in the same 
-iteration. They also use incremental search to avoid comparing points that have
-been compared in previous iterations. Also, sampled is used to lower the 
-overhead of local joins. Finally, random projection trees are used to 
-initialize the graph.
+a neighbor is likely to be a neighbor}. The idea is, given an approximation 
+graph of the nearest neighbor, then the approximation can be improved, 
+in an iterative fashion, by exploring the neighbors of the neighbors of each 
+point. To improve the running time, the authors use local join to avoid 
+comparing same points many times in the same iteration. They also use 
+incremental search to avoid comparing points that have been compared in 
+previous iterations. Also, sampled is used to lower the overhead of local 
+joins. Finally, random projection trees are used to initialize the graph.
 

### 2024-04-10 src/data.py
Create separate clkasses from train and test data

@@ -18,7 +18,7 @@ def get_dataset(
     return x, y
 
 
-class Data:
+class TrainData:
     """
     A class to store a dataset, and returns some info:
     - get sliding windows of the data
@@ -29,13 +29,13 @@ class Data:
 
     def __init__(self, dataset_name, window_size):
         x, y = get_dataset(dataset_name=dataset_name, train=True)
-        self.x_train: np.ndarray = x
-        self.y_train: np.ndarray = y
-        self.n_ts = self.x_train.shape[0]
-        self.ts_length = self.x_train.shape[1]
+        self.x: np.ndarray = x
+        self.y: np.ndarray = y
+        self.n_ts = self.x.shape[0]
+        self.ts_length = self.x.shape[1]
         self.window_size: int = window_size
         self.windows = sliding_window_view(
-            self.x_train, window_shape=self.window_size, axis=1
+            self.x, window_shape=self.window_size, axis=1
         ).reshape(-1, self.window_size)
         self.windows = StandardScaler().fit_transform(self.windows.T).T
         self._ts_covered: dict
@@ -46,7 +46,7 @@ class Data:
     def get_window_label(self, window_id):
         number_of_windows_per_ts = self.ts_length - self.window_size + 1
         ts_id = window_id // number_of_windows_per_ts
-        label = self.y_train[ts_id]
+        label = self.y[ts_id]
         if label in self._ts_covered:
             set_ts = self._ts_covered[label]
             set_ts.add(ts_id)
@@ -60,8 +60,52 @@ class Data:
         windows_labels = [self.get_window_label(wid) for wid in windows_ids]
         return np.array(windows_labels), self._ts_covered
 
-    def shapelet_transform(self, windows_ids):
-        candidates = self.windows[windows_ids]
-        windows = self.windows.reshape(self.n_ts, -1, self.window_size)
+
+class TestData:
+    def __init__(self, dataset_name, window_size):
+        x, y = get_dataset(dataset_name=dataset_name, train=False)
+        self.x: np.ndarray = x
+        self.y: np.ndarray = y
+        self.n_ts = self.x.shape[0]
+        self.ts_length = self.x.shape[1]
+        self.windows = sliding_window_view(
+            self.x, window_shape=window_size, axis=1
+        ).reshape(-1, window_size)
+        self.windows = StandardScaler().fit_transform(self.windows.T).T
+
+
+class Data:
+    """
+    A class to store a dataset, and returns some info:
+    - get sliding windows of the data
+    - return the label of a window given its index
[…]

### 2024-04-10 src/data.py
Refacoring and  fixing some minor bugs

@@ -87,6 +87,9 @@ class Data:
         self._train = TrainData(dataset_name, window_size)
         self._test = TestData(dataset_name, window_size)
 
+    def unique_labels(self):
+        return np.unique(self._train.y)
+
     def get_sliding_windows(self):
         return self._train.windows
 

### 2024-04-10 src/selection/info.py

@@ -52,7 +52,7 @@ class ClusterInfo:
             self.label,
             self.popularity * 100,
             len(self.ts_covered),
-            sum(data.y_train == self.label),
+            sum(data._train.y== self.label),
         ]
 
 
@@ -106,7 +106,7 @@ class ClustersInfo:
         return self.info_df
 
     def get_class_labels(self):
-        return np.unique(self.data.y_train)
+        return self.data.unique_labels()
 
     def get_clusters_of_labels(self, label, top=3):
         return self.info_df[self.info_df.label == label].id.values[:top]

### 2024-04-10 src/experiment.py
Add clustering tot he experiment

@@ -1,14 +1,19 @@
 from data import Data
 from selection import LIST_OF_ALGORITHMS, ClustersInfo
+from selection.methods import ClusterBased
 
 
 class Experiment:
     def __init__(self, description, data, algorithm) -> None:
         self.description = description
         self.selection_method = ClustersInfo(data, algorithm)
+        self.selection = ClusterBased(self.selection_method)
 
-    def run(self):
+    def info_df(self):
         return self.selection_method.info()
+
+    def get_shapelets(self):
+        self.selection.select()
 
     def __repr__(self) -> str:
         return self.description

### 2024-04-10 src/selection/info.py
refactor

@@ -18,14 +18,15 @@ class ClusterInfo:
     cluster_size: int
         the number of points assigned to this cluster
     label: int
-        The label assigned to this cluster based on the windows 
+        The label assigned to this cluster based on the windows
         contained in the cluster
     popularity: float
         The fraction of the windows representing the majority class (label)
     ts_covered: set
-        Set of time series of class label, where subsequences have 
+        Set of time series of class label, where subsequences have
         been assigned to this cluster
     """
+
     def __init__(self, cluster_id, data, windows_indices) -> None:
         self.cluster_id: int = cluster_id
         self.w_indices = windows_indices
@@ -52,7 +53,7 @@ class ClusterInfo:
             self.label,
             self.popularity * 100,
             len(self.ts_covered),
-            sum(data._train.y== self.label),
+            sum(data._train.y == self.label),
         ]
 
 
@@ -73,6 +74,7 @@ class ClustersInfo:
     cluster_info: DataFrame
         A DataFrame containing information about the clusters
     """
+
     def __init__(self, data, algorithm) -> None:
         self.data: Data = data
         self.algorithm: ClusterAlgorithm = algorithm
@@ -125,3 +127,4 @@ class ClustersInfo:
             if count == top:
                 break
         return np.array(selected_windows)
+

### 2024-04-10 src/experiment.py
Classification and evaluation starter code

@@ -1,19 +1,34 @@
 from data import Data
 from selection import LIST_OF_ALGORITHMS, ClustersInfo
 from selection.methods import ClusterBased
+from sklearn.linear_model import LogisticRegression
 
 
 class Experiment:
+    # TODO: Parameters have to come from some experiment metadata class!
     def __init__(self, description, data, algorithm) -> None:
         self.description = description
-        self.selection_method = ClustersInfo(data, algorithm)
+        self.data: Data = data
+        self.selection_method = ClustersInfo(self.data, algorithm)
+        # TODO: Initialize the selection mathod and selection somewhere else
         self.selection = ClusterBased(self.selection_method)
+        # TODO: Initialize classifier somewhere else and pass it to this class
+        self.classifier = LogisticRegression(random_state=0)
 
     def info_df(self):
         return self.selection_method.info()
 
     def get_shapelets(self):
-        self.selection.select()
+        return self.selection.select()
+
+    def classify(self):
+        shapelets = self.get_shapelets()
+        train = self.data.shapelet_transform(shapelets)
+        test = self.data.shapelet_transform(shapelets, train=False)
+        self.classifier.fit(train, self.data._train.y)
+        preds = self.classifier.predict(test)
+        # TODO: Separate the evaluation 
+        return sum(preds == self.data._test.y) / len(preds)
 
     def __repr__(self) -> str:
         return self.description

### 2024-04-11 Docs/paper/main.tex
Draft Introduction

\documentclass[sigconf, nonacm]{acmart}

\begin{document}
\title{Efficient Selection of Time Series Shapelets}

\author{Adam Charane}
\affiliation{%
	\institution{Free University of Bozen-Bolzano}
	\city{Bolzano}
	\state{Italy}
}
\email{acharane@unibz.it}

\author{Matteo Ceccarello}
\affiliation{%
	\institution{University of Padova}
	\city{Padova}
	\state{Italy}
}
\email{matteo.ceccarello@dei.unipd.it}

\author{Johann Gamper}
\affiliation{%
	\institution{Free University of Bozen-Bolzano}
	\city{Bolzano}
	\state{Italy}
}
\email{johann.gamper@unibz.it}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
	Praesent imperdiet, lacus nec varius placerat, est ex eleifend justo, a vulputate leo massa consectetur nunc. Donec posuere in mi ut tempus. Pellentesque sem odio, faucibus non mi in, laoreet maximus arcu. In hac habitasse platea dictumst. Nunc euismod neque eu urna accumsan, vitae vehicula metus tincidunt. Maecenas congue tortor nec varius pellentesque. Pellentesque bibendum libero ac dignissim euismod. Aliquam justo ante, pretium vel mollis sed, consectetur accumsan nibh. Nulla sit amet sollicitudin est. Etiam ullamcorper diam a sapien lacinia faucibus.
\end{abstract}

\maketitle

\section{Introduction}
Time series classification is similar to traditional classification where the
difference is that the order of the values of the time series is crucial.
The order holds discriminatory features for the time series. The literature
is rich of algorithms for classification, and one way to group these
algorithms is based on the techniques that each algorithm uses to exploit the
discriminatory features suggested by Bagnall et al.~\cite{bake_off}:
\begin{itemize}
	\item Distance based: Classification is based on using distance measures to
	      the raw time series.
	\item Feature based: Global features are extracted and the data is transformed
	      from the time domain to the feature domain spanned by the extracted
	      features, and the classification is done using a standard classifier.
	\item Interval based: Techniques in this family extract one or more intervals,
	      then statistical summaries of these intervals are used as features.
	\item Shapelet based: This technique exploits discriminatory patterns that
	      appear in time series of the same class.
	\item Dictionary based: These algorithms use as features the count of
	      repeating patterns.
	\item Convolution based: Features used are extracted from time series using
	      convolutions and pooling operations.
	\item Deep learning: Classification is done by feeding the data to neural
	      networks
	\item Hybrid: Algorithms in this category combine two or more approaches from
	      the previous techniques.
\end{itemize}
Our focus in this paper is on extracting discriminating patterns that represent
time series from each class. The extracted patterns then can be used by
algorithms from the shapelet based technique in order to execute the
classification task.

Time series classification using shapelets is usually done over three phases:
\begin{enumerate}
	\item Extraction of candidates~\cite{keogh_shapelet, random_shapelets, fss}:
	      Select subsequences that can be used as discriminating patterns.
	\item Evaluation of the candidates~\cite{alternative_measures,
		      shapelet_transform, silhoutte_shapelets}: Compute some statistics
	      about the extracted patterns in order to rank and select the best
	      ones (the most discriminating).
	\item Data Transformation~\cite{shapelet_transform}: After the best
	      candidates are chosen, the time series have to be transformed to a 
        format that can be used by standard classifiers.
\end{enumerate}
The extraction step can be done in different ways such as considering all the
subsequences. However, this results in high number of candidates. Namely, given
a dataset of $m$ time series each having a length $n$, then the number of
candidates considered is in the order of $\mathcal{O}(n^2m)$. Another popular
alternative is to randomly select subsequences from different time series and
from different positions. Another new approach is based on selecting
subsequences based on some statistics that can be computed efficiently such as
rolling mean, variance, change in the slope or other simple statistics.

Extracting candidates by itself can be done efficiently, but the challenge is
the second step, as the evaluation is based on the similarity between the
candidates, or between the candidates and the time series. For instance, if the
candidates are extracted using brute force(considering all the subsequences)
then computing evaluation statistics would require computing distances between
all pairs which is in the order of $O(n^4m^2)$. Clearly, this is very
inefficient and becomes impossible for even relatively small datasets.

Instead of evaluating by giving a score to each candidate independently, our
approach use efficient similarity search indexing approaches and exploits the
similarity structure in order to select the best candidates. The idea is that
good shapelets representing a class $C$ should have a lot nearest neighbors
that are from the same class, and at the same time, should have few or no
neighbors from other classes.

\section{Related work}
NN-Descent~\cite{NN-Descent} is based on the principle: \textit{A neighbor of
	a neighbor is likely to be a neighbor}. The idea is, given an approximation
graph of the nearest neighbor, then the approximation can be improved,
in an iterative fashion, by exploring the neighbors of the neighbors of each
point. To improve the running time, the authors use local join to avoid
comparing same points many times in the same iteration. They also use
incremental search to avoid comparing points that have been compared in
previous iterations. Also, sampled is used to lower the overhead of local
joins. Finally, random projection trees are used to initialize the graph.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}
\endinput

### 2024-04-12 Docs/paper/main.tex
todos

@@
 \end{itemize}
-Our focus in this paper is on extracting discriminating patterns that represent
-time series from each class. The extracted patterns then can be used by
-algorithms from the shapelet based technique in order to execute the
+Our focus in this paper is on extracting discriminating patterns, i.e. shapelets
+that represent time series from each class. The extracted patterns then can be 
+used by algorithms from the shapelet based technique in order to execute the
 classification task.
@@
 
+\section{Preliminaries}
+
 \section{Related work}
-NN-Descent~\cite{NN-Descent} is based on the principle: \textit{A neighbor of
-	a neighbor is likely to be a neighbor}. The idea is, given an approximation
-graph of the nearest neighbor, then the approximation can be improved,
-in an iterative fashion, by exploring the neighbors of the neighbors of each
-point. To improve the running time, the authors use local join to avoid
-comparing same points many times in the same iteration. They also use
-incremental search to avoid comparing points that have been compared in
-previous iterations. Also, sampled is used to lower the overhead of local
-joins. Finally, random projection trees are used to initialize the graph.
+This~\cite{keogh_shapelet} is where it all started, it has to be here.
+This paper~\cite{random_shapelets} uses random shapelet extraction.
+This paper~\cite{fss} uses neat tricks select shapelets.
+This paper~\cite{learning_shapelets} takes an alternative approach, where 
+instead of selecting shapelets it \emph{learns} non existing shapelets.
+We should mention the evaluation measures~\cite{alternative_measures}.
+We will use Silhouette~\cite{silhouettes} as motivation to exploit neighborhood
+structure.
+
+%NN-Descent~\cite{NN-Descent} is based on the principle: \textit{A neighbor of
+	%a neighbor is likely to be a neighbor}. The idea is, given an approximation
+%graph of the nearest neighbor, then the approximation can be improved,
+%in an iterative fashion, by exploring the neighbors of the neighbors of each
+%point. To improve the running time, the authors use local join to avoid
+%comparing same points many times in the same iteration. They also use
+%incremental search to avoid comparing points that have been compared in
+%previous iterations. Also, sampled is used to lower the overhead of local
+%joins. Finally, random projection trees are used to initialize the graph.
 

### 2024-04-12 Docs/paper/main.tex
Drafting Background and related work

@@
 Time series classification is similar to traditional classification where the
-difference is that the order of the values of the time series is crucial.
-The order holds discriminatory features for the time series. The literature
-is rich of algorithms for classification, and one way to group these
-algorithms is based on the techniques that each algorithm uses to exploit the
-discriminatory features suggested by Bagnall et al.~\cite{bake_off}:
+difference is that the order of the values of the time series is crucial, since 
+it holds discriminatory features of the time series. The literature is rich of
+algorithms for classifying time series, and one way to group them,
+suggested by Bagnall et al.~\cite{bake_off}, is based on the techniques that
+each algorithm uses to exploit the discriminatory features:
 \begin{itemize}
@@
 similarity structure in order to select the best candidates. The idea is that
-good shapelets representing a class $C$ should have a lot nearest neighbors
+good shapelets representing a class $C$ should have a lot of nearest neighbors
 that are from the same class, and at the same time, should have few or no
@@
 
-\section{Preliminaries}
+\section{Background \& Related work}
+In a classification setup, the idea of a shapelet, introduced by Ye and 
+Keogh~\cite{keogh_shapelet}, is that time series from the same class must have 
+some shared pattern that is similar to all of them and less similar to time 
+series from other classes. A pattern is simply a subsequence of the time series,
+and the (dis-)similarity is a distance measure, usually the normalized Euclidean
+distance. Besides the high accuracy of shapelets, they are also interpretable,
+since the subsequences chosen as shapelets can be interpreted as the event that
+happened at that period of time.
+Although the idea being simple, it introduced many challenges. The two 
+challenges we are concerned about are:
+\begin{enumerate}
+  \item How to choose patterns?
+  \item How to evaluate them?
+\end{enumerate}
+In the original work~\cite{keogh_shapelet}, the authors considered all the 
+subsequences as patterns, and for the evaluation part, they ranked the patterns
+by assigning to each a score. The score was the information gain, which is based
+on the Shanon entropy, before and after considering the subsequence. At the end,
+only the subsequences with the highest score were selected as shapelets.
+Despite that only few patterns were selected, they all had to be evaluated,
+which means, the similarity between all of them had to be computed in order to 
+assign a score. Computing the similarity was the bottleneck of the computation,
+in order to speed it up, the authors had to apply early computation pruning.
+The pruning resulted in a speed up, but still the approach is very slow,
+and not practical for medium and large size datasets. To improve the running 
+time, random selection of shapelets~\cite{random_shapelets} have been 
+introduced. The authors empirically shown that by selecting a large number of
+shapelets, i.e., by randomly selecting subsequences with different lengths from 
+random time series at random positions, the computation becomes much faster
+since only few patterns are considered, while the accuracy remains competitive.
 
-\section{Related work}
-This~\cite{keogh_shapelet} is where it all started, it has to be here.
-This paper~\cite{random_shapelets} uses random shapelet extraction.
-This paper~\cite{fss} uses neat tricks select shapelets.
+In order to speed up the selection process, Ji et al.~\cite{fss} introduced 
+Fast Shapelet Transform (FSS), that works on two different steps. First, they 
+sample  time series with the help of subclass splitting. This reduces the 
+number of time series to be considered significantly, without affecting the 
+accuracy, since the goal is to look for similar patterns and not all patterns.
+Second, the Local Farthest Deviation Points (LFDP) are selected and used as the
+endpoints of the subsequences. An LFDP is the point that has the largest 
+distance from the subsequence fitting line. The intuition is that the constant
+subsequences should be ignored and only subsequences where there are changes
+should be interesting.
+
+Regardless of the technique used in order to select the candidates, the 
+evaluation in the above approaches was done using the information gain.
+Lines et al.~\cite{alternative_measures} were the first to consider using
+different evaluation methods instead of the information gain. They considered
+as evaluation measures statistical tests to verify if points from the same
+distribution or not, such as Kruskal Wallis and Mood's medians. On another 
+the same authors use the ANOVA F-statistic~\cite{shapelet_transform} as an 
+evaluation measure, which was not significantly better than the information 
+gain, but on average, shapelets selected using the F-statistic resulted in high
+classification accuracy.
+
 This paper~\cite{learning_shapelets} takes an alternative approach, where 
@@
 structure.
+
+\section{Preliminaries}
+
+\section{Selection}
 

### 2024-04-16 Docs/paper/main.tex
Preliminaries

@@
 classification accuracy.
+A recent work~\cite{silhoutte_shapelets} proposed the usage of Silhouettes
+\cite{silhouettes} as an evaluation measure, the results were competitive with
+the F-statistic and information gain, but the silhouette had much better results
+when the number of shapelets used to classify was small.
 
@@
 \section{Preliminaries}
+This section defines the key concepts used in the paper.
+\begin{definition}[Time Series]
+	A time series $T$ is sequence of $n$ consecutive real numbers:
+	$T = [T_1, T_2, \dots, T_n]$.
+\end{definition}
 
-\section{Selection}
+\begin{definition}[Subsequence]
+	A subsequence of length $l$ from a time series $T$ starting from position $i$
+	is an ordered set of consecutive values from $T$ denoted as:
+	$S_{i, l} = [T_i, T_{i+1},, \dots, T_{i+l-1}]$.
+\end{definition}
+
+\begin{definition}[Sliding window]
+	A sliding window of length $l$ over a time series $T$ is the set of
+	all subsequences of length $l$ extracted from $T$:
+	$W_T^l = [S_{1, l}, S_{2, l}, \dots S_{n-l+1, l}]$.
+\end{definition}
+
+A dataset consists of $m$ time series, each having a class label $c \in [1, 2,
+		\dots, K]$. The sliding windows $W_T^l$ all have the same class label $c$ of the
+time series $T$ they were extracted from. To compute the evaluation score of
+each window using the Silhouette, we need first to introduce a (dis-)similarity
+measure. In this work, we consider the normalized Euclidean distance:
+
+\begin{definition}[Normalized Euclidean distance]
+	The normalized Euclidean distance between two subsequences $S$ and $Q$ of
+	length $l$:
+	$$d(S, Q) = \sqrt{\sum_{i=1}^l (\frac{S_i - \mu_S}{\sigma_S} - \frac{Q_i -
+				\mu_Q}{\sigma_Q})^2}$$
+	where $\mu_S$ and $\sigma_S$ represent the mean and standard deviation of
+	$S$ respectively.
+\end{definition}
+
+Finally, we define the silhouette score for a subsequence:
+
+\begin{definition}[Silhouette score of a subsequence]
+	Given a set $\mathcal{D}$ of subsequences (sliding windows) and their
+	corresponding labels, the silhouette score for a subsequence $S_i$ having label
+	$c$ is:
+	$$
+		s(S_i) = \frac{b - a}{\max{(a, b)}}
+	$$
+	where:
+	\begin{itemize}
+		\item $b = \frac{\sum_{j \in I \setminus I_c}d(S_i, S_j)}{|I \setminus I_c|}$,
+		\item $a = \frac{\sum_{k \in I_c}d(S_i, S_k)}{|I_c|}$,
+		\item $I$ is the set of indices of all subsequences in $\mathcal{D}$, and
+		\item $I_c$ is the set of indices of subsequences having label $c$.
+	\end{itemize}
+\end{definition}
+
+Notice, that our definition of the silhouette, the term $b$ considers all 
+subsequences from other classes instead of only considering only the class that
+is closest to the subsequence $S_i$.
 

### 2024-04-16 Docs/paper/main.tex
Drafting the approach

@@
 	\begin{itemize}
+		\item $a = \frac{\sum_{k \in I_c}d(S_i, S_k)}{|I_c|}$,
 		\item $b = \frac{\sum_{j \in I \setminus I_c}d(S_i, S_j)}{|I \setminus I_c|}$,
-		\item $a = \frac{\sum_{k \in I_c}d(S_i, S_k)}{|I_c|}$,
 		\item $I$ is the set of indices of all subsequences in $\mathcal{D}$, and
@@
 
-Notice, that our definition of the silhouette, the term $b$ considers all 
+Notice that our definition of the silhouette, the term $b$ considers all
 subsequences from other classes instead of only considering only the class that
@@
 
-%NN-Descent~\cite{NN-Descent} is based on the principle: \textit{A neighbor of
-%a neighbor is likely to be a neighbor}. The idea is, given an approximation
-%graph of the nearest neighbor, then the approximation can be improved,
-%in an iterative fashion, by exploring the neighbors of the neighbors of each
-%point. To improve the running time, the authors use local join to avoid
-%comparing same points many times in the same iteration. They also use
-%incremental search to avoid comparing points that have been compared in
-%previous iterations. Also, sampled is used to lower the overhead of local
-%joins. Finally, random projection trees are used to initialize the graph.
+\section{Approach}
+The bottleneck of the computation is in the evaluation part, where, in order to
+compute the Silhouette score (also for the F-statistic and all other measures),
+all pairs of distances have to be computed. To avoid that, we will use the fact
+that shapelets that have a high silhouette score are of good quality for time
+series classification. Since we know, in a clustering setup, silhouette is high
+when he clusters are compact and far from each other, then the problem of
+finding good quality shapelets reduces to finding clusters with high silhouette
+scores.
+
+We start by extracting the set of sliding windows $W$ from all the time series
+$W = \bigcup_T W_T^l$, then we cluster all the windows in $W$.
+For each cluster $C_i$, we assign a label $c$ using a majority vote, i.e., the
+label of cluster $C_i$ is the label of the set with largest cardinal, consisting
+of windows assigned to cluster $C_i$ that have the same class label.
+
+\section{Experimental Evaluation}
+
+\section{Conclusion}
 

### 2024-04-17 Docs/paper/main.tex
Learning shapelets

@@
 
-This paper~\cite{learning_shapelets} takes an alternative approach, where
-instead of selecting shapelets it \emph{learns} non existing shapelets.
-We should mention the evaluation measures~\cite{alternative_measures}.
-We will use Silhouette~\cite{silhouettes} as motivation to exploit neighborhood
-structure.
+While all the previous work found shapelets by extracting subsequences from the
+time series, Grabocka et al.~\cite{learning_shapelets} take an alternative
+approach, where instead of selecting shapelets, they \emph{learn} them by
+solving an optimization problem.
 

### 2024-04-17 Docs/paper/main.tex
Add clustering example

@@
 \email{johann.gamper@unibz.it}
+
+\newcommand{\mc}[1]{\textcolor{teal}{ --- MC: #1 --- }}
+\newcommand{\ac}[1]{\textcolor{orange}{ --- AC: #1 --- }}
+\newcommand{\jg}[1]{\textcolor{red}{ --- JG: #1 --- }}
 
@@
 subsequences from other classes instead of only considering only the class that
-is closest to the subsequence $S_i$.
+is closest to the subsequence $S_i$. Next, a fixed number of subsequences from
+each class are selected as shapelets, based on the highest score achieved.
 
@@
 The bottleneck of the computation is in the evaluation part, where, in order to
-compute the Silhouette score (also for the F-statistic and all other measures),
+compute the Silhouette score (same for the F-statistic and the other measures),
 all pairs of distances have to be computed. To avoid that, we will use the fact
-that shapelets that have a high silhouette score are of good quality for time
-series classification. Since we know, in a clustering setup, silhouette is high
-when he clusters are compact and far from each other, then the problem of
-finding good quality shapelets reduces to finding clusters with high silhouette
-scores.
+that shapelets with a high silhouette score are of good quality for time
+series classification. Since in a clustering setup, silhouette index of a 
+cluster is high when the cluster is compact and far from other clusters, and 
+the silhouette score of a cluster is simply the average of the silhouette 
+scores of the subsequences in that cluster, then the subsequences with high 
+silhouette are contained in the cluster.
 
+\subsection{Clustering-based shapelets}
+\begin{figure*}
+  \includegraphics[width=\textwidth]{imgs/clustering.png}
+  \caption{Sliding windows of length 40 from \emph{CBF} dataset, with centers
+  (red dots) of 100 clusters using K-Means algorithm.\label{fig:clustering}}
+\end{figure*}
 We start by extracting the set of sliding windows $W$ from all the time series
@@
 of windows assigned to cluster $C_i$ that have the same class label.
+Figure~\ref{fig:clustering} shows an example from the \emph{CBF} dataset plotted
+on the plan by keeping only the first two PCA components. The gray dots in the
+left sub-figure are the windows of length 40, and the red dots are the centers
+of the clusters obtained y running K-Means algorithm with 100 clusters. The
+right sub-figure shows the labels of the subsequences.
 

### 2024-04-17 src/data.py
add parameters for optional preprocessing steps

@@ -27,7 +27,7 @@ class TrainData:
     - Given a list of windows indices return the shapelet transform
     """
 
-    def __init__(self, dataset_name, window_size):
+    def __init__(self, dataset_name, window_size, preprocessors=None):
         x, y = get_dataset(dataset_name=dataset_name, train=True)
         self.x: np.ndarray = x
         self.y: np.ndarray = y
@@ -38,6 +38,9 @@ class TrainData:
             self.x, window_shape=self.window_size, axis=1
         ).reshape(-1, self.window_size)
         self.windows = StandardScaler().fit_transform(self.windows.T).T
+        if preprocessors:
+            for processor in preprocessors:
+                self.windows = processor.fit_transform(self.windows)
         self._ts_covered: dict
 
     def get_sliding_windows(self):
@@ -83,8 +86,8 @@ class Data:
     - Given a list of windows indices return the shapelet transform
     """
 
-    def __init__(self, dataset_name, window_size):
-        self._train = TrainData(dataset_name, window_size)
+    def __init__(self, dataset_name, window_size, preprocessors=None):
+        self._train = TrainData(dataset_name, window_size, preprocessors)
         self._test = TestData(dataset_name, window_size)
 
     def unique_labels(self):

### 2024-04-19 Docs/paper/main.tex
Clustering based shapelets, example and some questions

@@
 
-\author{Adam Charane}
-\affiliation{%
-	\institution{Free University of Bozen-Bolzano}
-	\city{Bolzano}
-	\state{Italy}
-}
-\email{acharane@unibz.it}
+%\author{Adam Charane}
+%\affiliation{%
+%\institution{Free University of Bozen-Bolzano}
+%\city{Bolzano}
+%\state{Italy}
+%}
+%\email{acharane@unibz.it}
 
-\author{Matteo Ceccarello}
-\affiliation{%
-	\institution{University of Padova}
-	\city{Padova}
-	\state{Italy}
-}
-\email{matteo.ceccarello@dei.unipd.it}
+%\author{Matteo Ceccarello}
+%\affiliation{%
+%\institution{University of Padova}
+%\city{Padova}
+%\state{Italy}
+%}
+%\email{matteo.ceccarello@dei.unipd.it}
 
-\author{Johann Gamper}
-\affiliation{%
-	\institution{Free University of Bozen-Bolzano}
-	\city{Bolzano}
-	\state{Italy}
-}
-\email{johann.gamper@unibz.it}
+%\author{Johann Gamper}
+%\affiliation{%
+%\institution{Free University of Bozen-Bolzano}
+%\city{Bolzano}
+%\state{Italy}
+%}
+%\email{johann.gamper@unibz.it}
 
@@
 $W = \bigcup_T W_T^l$, then we cluster all the windows in $W$.
-For each cluster $C_i$, we assign a label $c$ using a majority vote, i.e., the
-label of cluster $C_i$ is the label of the set with largest cardinal, consisting
-of windows assigned to cluster $C_i$ that have the same class label.
-Figure~\ref{fig:clustering} shows an example from the \emph{CBF} dataset plotted
-on the plan by keeping only the first two PCA components. The gray dots in the
-left sub-figure are the windows of length 40, and the red dots are the centers
-of the clusters obtained y running K-Means algorithm with 100 clusters. The
-right sub-figure shows the labels of the subsequences.
+For each cluster $C_i$, we count the fraction of windows from each class
+(popularity), and only keep clusters where the popularity of one class is beyond
+a threshold. Figure~\ref{fig:clustering} shows an example from the \emph{CBF}
+dataset plotted on the plan by keeping only the first two PCA components. The
+gray dots in the left sub-figure are the windows of length 40, and the red dots
+are the centers of the clusters obtained y running K-Means algorithm with 100
+clusters. The right sub-figure shows the labels of the subsequences.
+Figure~\ref{fig:clustering_threshold} shows the same windows, but this time
+after removing all the clusters where the popularity of the dominant class is
+below some certain threshold.
+The idea is that we want to get rid of all subsequences from different classes
+that are similar to each other, since a shapelet should be a pattern repeated
+between time series from the same class.
+\begin{figure*}
+	\includegraphics[width=\textwidth]{imgs/clustering_thresholds.png}
+	\caption{Windows left after removing all the clusters where the popularity of
+		the dominant class is below a threshold\label{fig:clustering_threshold}.}
+\end{figure*}
+
+\ac{
+	There are many questions/decisions that have to be made:
+	\begin{itemize}
+		\item How many clusters we should use?
+		\item Which clusters to keep (so far I am using the popularity)?
+		\item Should we use all the clusters kept?
+	\end{itemize}
+	Some information that we have and that have not been used yet:
+	\begin{itemize}
+		\item In each cluster, besides the popularity of each class, we know also
+		      how many time series are covered. For example, in the \emph{CBF}
+		      dataset, we have 30 training examples, 10 for each class. If we have
+		      two clusters with popularity, let's say 90\% for class 1, the windows
+		      from cluster 1 are all extracted from 2 time series, and in the second
+		      clusters, the windows are from all 10 time series, clearly we have to
+		      favor the second cluster, since it has a pattern that is repeated
+		      between all the of the time series from the same class. How can we
+		      use this information?
+		\item How to use the distance between the centers of the clusters?
+	\end{itemize}
+}
 

### 2024-04-19 Docs/paper/main.tex
Similarity search indexing for shapelets general idea

@@
 
+\subsection{Similarity search indexing-based shapelets}
+In the previous subsection, we have used clustering to get rid of the evaluation
+part and reduce the computational overhead. However, we used clustering because
+it is the first intuitive step that comes in order to optimize for the 
+silhouette score.
+Since what we care about is the structure of similarity search, i.e., we want to
+know, for each subsequence, which other subsequences are the most similar. To
+answer these questions, one can use efficient indexing data structures from the
+nearest-neighbor problem literature, which is more efficient compared to 
+clustering in terms of time complexity and running time.
+
 \section{Experimental Evaluation}

### 2024-04-19 Docs/paper/main.tex
Some remarks concerning clustering based shapelets

@@
 that are similar to each other, since a shapelet should be a pattern repeated
-between time series from the same class.
+between time series from the same class. Notice that the threshold might have a
+high impact on the results, since when we apply a $90\%$ threshold, we have lost
+some important clusters for class 2 (top right in Figure
+\ref{fig:clustering_threshold}, as well as some clusters for class 1 (in the 
+middle of the plot).
 \begin{figure*}
@@
 		\item Which clusters to keep (so far I am using the popularity)?
+    \item Another approach would be to select top clusters from each class. This
+      is especially relevant when a class is never dominating in terms of 
+      population, that is, if we put a threshold, there are no clusters
+      where a class $c$ has more than the threshold of the population, then we
+      will end up with no shapelets representing a specific class.
 		\item Should we use all the clusters kept?

### 2024-04-19 Docs/paper/main.tex
Some minor changes

@@
 In the previous subsection, we have used clustering to get rid of the evaluation
-part and reduce the computational overhead. However, we used clustering because
-it is the first intuitive step that comes in order to optimize for the 
-silhouette score.
+part and reduce the computational overhead. We used clustering because it is the
+first intuitive idea in order to optimize for the silhouette score.
 Since what we care about is the structure of similarity search, i.e., we want to
-know, for each subsequence, which other subsequences are the most similar. To
-answer these questions, one can use efficient indexing data structures from the
-nearest-neighbor problem literature, which is more efficient compared to 
-clustering in terms of time complexity and running time.
+know, for each subsequence, which other subsequences are the most similar. An
+alternative approach to answer these questions, one can use efficient indexing
+data structures from the nearest-neighbor problem literature, which is more 
+efficient compared to clustering in terms of time complexity and running time.
 

### 2024-04-22 Docs/paper/main.tex
Experimental evaluation general

@@
 \section{Experimental Evaluation}
+In this section we compare the running time and the accuracy of our two
+approaches with the baselines, namely:
+(1) Ultra fast shapelet transform, which randomly select shapelets, evaluate
+selects the ones with the highest rank. (2) Learning shapelets. (3) Fast them,
+and Shapelet Selection (FSS).
 

### 2024-04-23 src/data.py
Fix bug when _ts_covered is not itialized

@@ -41,12 +41,14 @@ class TrainData:
         if preprocessors:
             for processor in preprocessors:
                 self.windows = processor.fit_transform(self.windows)
-        self._ts_covered: dict
+        self._ts_covered: dict | None = None
 
     def get_sliding_windows(self):
         return self.windows
 
     def get_window_label(self, window_id):
+        if self._ts_covered is None:
+            self._ts_covered = {}
         number_of_windows_per_ts = self.ts_length - self.window_size + 1
         ts_id = window_id // number_of_windows_per_ts
         label = self.y[ts_id]
@@ -59,7 +61,6 @@ class TrainData:
         return label
 
     def windows_labels_and_covered_ts(self, windows_ids):
-        self._ts_covered = {}
         windows_labels = [self.get_window_label(wid) for wid in windows_ids]
         return np.array(windows_labels), self._ts_covered
 
